Kinematics of the Palomar 5 stellar stream from RR Lyrae stars

### 2019-09-30 paper/main.tex
fix changes

\begin{abstract}
% Context
Thin stellar streams, formed from the tidal disruption of globular clusters, are important gravitational tools, sensitive to both global and small-scale properties of dark matter.
The Palomar 5 stellar stream (Pal 5) is an exemplar stream within the Milky Way: Its $\sim 20\degr$ tidal tails connect back to the progenitor cluster, and the stream has been used to study the shape, total mass, and substructure fraction of the dark matter distribution of the Galaxy.
However, most details of the phase-space distribution of the stream are not fully explained, and dynamical models that use the stream for other inferences are therefore incomplete.
% Aims
Here we aim to measure distance and kinematic properties along the Pal 5 stream in order to motivate improved models of the system. %Pal 5 stream.
We use a large catalog of RR Lyrae-type stars (RRLs) with astrometric data from the \Gaia\ mission to probabilistically identify RRLs in the Pal 5 stream.
RRLs are useful because they are intrinsically-luminous standard candles and their distances can be inferred with small relative precision ($\sim3\%$).
% Methods
% Results
By building a probabilistic model of the Pal 5 \changes{cluster and stream} in proper motion and distance, we find \NRRL\ RRLs consistent with being members of the cluster (\NRRcl) and stream (\NRRtails).
Using these RRLs, we detect gradients in distance and proper motion along the stream, and provide an updated measurement of the distance to the Pal 5 cluster using the RRLs, $d = \clderr$.
% Conclusions
We provide a catalog of Pal 5 RRLs with inferred membership probabilities for future modeling work.
\end{abstract}

@@

We cross-match the full PS1 RRL catalog with \Gaia~DR2 \citep{Gaia:2018}, using a 1" sky separation tolerance, in order to retrieve astrometric information for these stars.
Although not all PS1 RRLs have been identified as RRLs from \Gaia\ photometry, all of the PS1 RRLs are present in the main point source catalog.
\figurename~\ref{fig:pm} demonstrates that Pal 5 cluster and stream members appear in the RRL catalog: This figure shows the \Gaia\ \DR{2} proper motions for all RRLs in the selected sky window (indicated in the caption) within the distance range $18 < d < 25~\kpc$, then the same for RRLs in a tight selection around the observed stream track (see \sectionname~\ref{sec:membership} for details), and then finally the same for RRLs \emph{excluding} the stream region.

>>>>

\begin{abstract}
% Context
Thin stellar streams, formed from the tidal disruption of globular clusters, are important gravitational tools, sensitive to both global and small-scale properties of dark matter.
The Palomar 5 stellar stream (Pal 5) is an exemplar stream within the Milky Way: Its $\sim 20\degr$ tidal tails connect back to the progenitor cluster, and the stream has been used to study the shape, total mass, and substructure fraction of the dark matter distribution of the Galaxy.
However, most details of the phase-space distribution of the stream are not fully explained, and dynamical models that use the stream for other inferences are therefore incomplete.
% Aims
Here we aim to measure distance and kinematic properties along the Pal 5 stream in order to motivate improved models of the system. %Pal 5 stream.
We use a large catalog of RR Lyrae-type stars (RRLs) with astrometric data from the \Gaia\ mission to probabilistically identify RRLs in the Pal 5 stream.
RRLs are useful because they are intrinsically-luminous standard candles and their distances can be inferred with small relative precision ($\sim3\%$).
% Methods
% Results
By building a probabilistic model of the Pal 5 cluster and stream in proper motion and distance, we find \NRRL\ RRLs consistent with being members of the cluster (\NRRcl) and stream (\NRRtails).
Using these RRLs, we detect gradients in distance and proper motion along the stream, and provide an updated measurement of the distance to the Pal 5 cluster using the RRLs, $d = \clderr$.
% Conclusions
We provide a catalog of Pal 5 RRLs with inferred membership probabilities for future modeling work.
\end{abstract}

@@

We cross-match the full PS1 RRL catalog with \Gaia~DR2 \citep{Gaia:2018}, using a 1" sky separation tolerance, in order to retrieve astrometric information for these stars.
Although not all PS1 RRLs have been identified as RRLs from \Gaia\ photometry, all of the PS1 RRLs are present in the main point source catalog.
\figurename~\ref{fig:pm} demonstrates that Pal 5 \changes{cluster and stream} members appear in the RRL catalog: This figure shows the \Gaia\ \DR{2} proper motions for all RRLs in the selected sky window (indicated in the caption) within the distance range $18 < d < 25~\kpc$, then the same for RRLs in a tight selection around the observed stream track (see \sectionname~\ref{sec:membership} for details), and then finally the same for RRLs \emph{excluding} the stream region.
